Exploratory Data Analysis for College Scorecard Dataset

### 2017-02-22 Draft-xuezhou-sjb.Rmd
further fix

###5.1.2 Student Body<a id="5.1.2"></a> 
####5.1.2.1 Percentage first-generation students<a id="5.1.2.1"></a>
This element identifies the share of federally aided students at the institution who report that they are the first generaion on the FAFSA (Free Application for Federal Student Aid). First generation students are defined as students whose parents didn't attend a college.

@@

  ggplot(data=data_clean , aes(PAR_ED_PCT_1STGEN)) + 
    geom_histogram(breaks=seq(0.1, 1, by =0.1), 
                   col="white", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for Percentage first-generation students") +
    labs(x="PAR_ED_PCT_1STGEN", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```

@@

With the following codes, we extracted the IV institutions with the highest and lowest average net price. Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.

@@

From the summary results, we learned that the maximum value of the variable is 98.56%(Capella University) while the minimum value is about 0.73% (University of Notre Dame). From the mean value, we can tell that in most colleges, around 56% of the students are financially independent. The standard deviation value 23.3% indicates a relatively high difference among different colleges, which can also be concluded from the difference between the maximum value and the minimum value. Below is the code to generate a visual representation in the histogram

@@

  #histogram for NPT_COM
  ggplot(data=data_clean , aes(DEP_STAT_PCT_IND)) + 
    geom_histogram(breaks=seq(0.1, 1, by =0.1), 
                   col="white", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for Percentage of students who are financially independent") +
    labs(x="DEP_STAT_PCT_IND", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```

@@

This element, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the financial aid an institution provides to low-income students. We are using the code below to summarize this variable and extract the institutions with the highest and lowest value.

@@

  ggplot(data=data_clean , aes(PCTPELL)) + 
    geom_histogram(breaks=seq(0, 1, by =0.1), 
                   col="white", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for Percentage of undergraduates who receive a Pell Grant") +
    labs(x="PCTPELL", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```

@@

####5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan<a id="5.1.4.2"></a>

This element, as reported in the IPEDs Student Finaicial Aid component, shows the share of undergraduate students who received federal loans in a given year. It can provide important context to figures related to debt, repayment and non-repayment.We are using the code below to summarize this variable and extract the institutions with highest and lowest value.

@@

  ggplot(data=data_clean , aes(PCTFLOAN)) + 
    geom_histogram(breaks=seq(0, 1, by =0.1), 
                   col="white", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for Percent of all federal undergraduate students receiving a federal student loan") +
    labs(x="PCTFLOAN", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```

@@

This variable recorded debt for students who have completed the education. We are using the code below to summarize this variable and extracted the institutions with largest and smallest value.

@@

#6. Relationship between variables<a id="6"></a>
##6.1. The relationship between average net cost for Title IV institutions and State<a id="6.1"></a>

To analyze the relationship between the average net cost for Title IV institutions and State, we grouped the data into the state level to take a further look. We would like to investigate the distribution of institutions ' average net price in each state.

@@

To compare the percentage of Pell Grant in different control types, we grouped the Percentage of Pell Grant data using college type to take a further analysis.

@@

To compare the percentage of first-generation students in each states, we grouped data into the state level to take a further analysis.

@@

##6.5. The relationship between family income and percentage of Pell Grant<a id="6.5"></a>

@@

To compare the percentage of financially-independent students in the institutions with different college type, we grouped the data by the control type(private for-profit, public, private nonprofit) to take a further analysis.

@@

To compare the percentage of students who receive the federal loans in the given year in each schools control type, we grouped the data by the control type(private for-profit, public, private non-profit) to take a further analysis. A stacked histogram is also used to generate a more visual presentation.

@@

#6.13 The Relashionship between Control and Median earnings of Students Working and Not Enrolled 6 Years after Entry<a id="6.13"></a>

>>>>

###5.1.2 Student Body<a id="5.1.2"></a> 
####5.1.2.1 Percentage of First-generation Students<a id="5.1.2.1"></a>
This element identifies the share of federally aided students at the institution who report that they are the first generaion on the FAFSA (Free Application for Federal Student Aid). First generation students are defined as students whose parents didn't attend a college.

@@

  ggplot(data=data_clean , aes(PAR_ED_PCT_1STGEN)) + 
    geom_histogram(breaks=seq(0.1, 1, by =0.1), 
                   col="white", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for Percentage First-Generation Students") +
    labs(x="PAR_ED_PCT_1STGEN", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```

@@

With the following codes, we extracted the IV institutions' basic demographic information with the highest and lowest average net price. Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.

@@

From the summary results, we learned that the maximum value of the variable is 98.56% (Capella University) while the minimum value is about 0.73% (University of Notre Dame). From the mean value, we can tell that in most colleges, around 56% of the students are financially independent. The standard deviation value 23.3% indicates a relatively high difference among different colleges, which can also be concluded from the difference between the maximum value and the minimum value. Below is the code to generate a visual representation in the histogram

@@

  #histogram for NPT_COM
  ggplot(data=data_clean , aes(DEP_STAT_PCT_IND)) + 
    geom_histogram(breaks=seq(0.1, 1, by =0.1), 
                   col="white", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for Percentage of Students Who Are Financially Independent") +
    labs(x="DEP_STAT_PCT_IND", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```

@@

This element, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the financial aid an institution provides to low-income students. We are using the code below to summarize this variable(min, max, mean, median and standard deviation) and extract the institutions with the highest and lowest value.

@@

  ggplot(data=data_clean , aes(PCTPELL)) + 
    geom_histogram(breaks=seq(0, 1, by =0.1), 
                   col="white", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for Percentage of Undergraduates Who Receive a Pell Grant") +
    labs(x="PCTPELL", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```

@@

####5.1.4.2 Percent of All Federal Undergraduate Students Receiving a Federal Student Loan<a id="5.1.4.2"></a>

This element, as reported in the IPEDs Student Finaicial Aid component, shows the share of undergraduate students who received federal loans in a given year. It can provide important context to figures related to debt, repayment and non-repayment. We were using the code below to summarize this variable(min, max, mean, median and standard deviation) and extracted the institutions with highest and lowest value.

@@

  ggplot(data=data_clean , aes(PCTFLOAN)) + 
    geom_histogram(breaks=seq(0, 1, by =0.1), 
                   col="white", 
                   aes(fill=..count..),na.rm=TRUE)+
    labs(title="Histogram for Percent of all Federal Undergraduate Students Receiving a Federal Student Loan") +
    labs(x="PCTFLOAN", y="Count") +
    theme(plot.title = element_text(hjust = 0.5))
```

@@

This variable recorded debt for students who have completed the education. We were using the code below to summarize this variable(min, max, mean, median and standard deviation) and extracted the institutions with largest and smallest value.

@@

#6. Relationship between variables<a id="6"></a>
##6.1. The Relationship between Average Net Cost for Title IV institutions and State<a id="6.1"></a>

To analyze the relationship between the average net cost for Title IV institutions and State, we grouped the data into state level to take a further analysis on each state's min, max and median value of the Average Net Cost - We would like to investigate the distribution of institutions ' average net price in each state.

@@

To compare the percentage of Pell Grant in different control types, we grouped the Percentage of Pell Grant data using college type to take a further analysis on each college type's min, max and median value of Percentage of Pell Grant.

@@

To compare the percentage of first-generation students in each states, we grouped data into the state level to take a further analysis on the min, max and median value of Percentage first-generation students on the state base. The top 5 states as well as the lowest 5 states are extracted.

@@

##6.5. The relationship between Family Income and Percentage of Pell Grant<a id="6.5"></a>

@@

To compare the percentage of financially-independent students in the institutions with different college type, we grouped the data by the control type(private for-profit, public, private nonprofit) to take a further analysis on the min, max and median value for each type.

@@

To compare the percentage of students who receive the federal loans in the given year in each schools control type, we grouped the data by the control type(private for-profit, public, private non-profit) to take a further analysis on the min, max and median value of the variable seperately. A stacked histogram is also used to generate a more visual presentation.

@@

##6.13 The Relashionship between Control and Median earnings of Students Working and Not Enrolled 6 Years after Entry<a id="6.13"></a>
